---
title: "Exam"
author: 'Group_1 : Aditi, Shwesin, Theo, Akash and Daanyaal'
date: "2022-09-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.



#As a group we have devided the tasks between each other
#through each others mistakes we have learnt
#The road is visible in script DYW_Exam_Script
#Now we have produced a final recipe to tidy this specific data
#---------------------- Refined script and steps -----------------------------#


**Packages**

-   [`Reference - ways to load and install packages`](https://statsandr.com/blog/an-efficient-way-to-install-and-load-r-packages/){.uri}

```{r Install and load packages - Alternative 1}
#List of packages
#install.packages(c(tidyverse, ggplot2, here, pacman))
library(tidyverse)
library(here)
library(ggplot2)
library(fs)
library(knitr)
#pacman::p_load(ggplot2, tidyverse, here, fs)

```



**Directory and Environment**

-   `Ctrl+Shift+A` to reformat code in the code chunks/R script file
-   `Ctrl+Shift+M` for `%>%`

```{r }
#getwd()  - current directory
#here()   - current directory
#setwd()  - change working directory
#ls()     - files in environment
#remove(ls()) - clears environment - OBS!
#dir()    - files in current directory
#dir_tree(here()) - draws a tree of the current directory
```

**File**

-   Loading file into the environment
-   Data manipulation

```{r Load file}
#Step 2 - Reading the text files, make sure directory is correct
#Load file into work environment 
#Separate columns using delimiter ("tab" - \t)
#Set working directory setwd() if things are not working

here()
setwd("C:/Users/asi006/Downloads/RMED/Advanced_Group_one_project/Data")
nontidy_data <- read_delim("exam_nontidy.txt", delim = "\t")
tidy_data <- nontidy_data
join_data <- read_delim("exam_joindata.txt", delim = "\t")


```


```{r Data wrangling}

#Step 4 - The Pipe setup of multiple tidy mutations and transfomation
#----------------Will yield to the tidiest data -------------------#

tidy_data <- tidy_data%>%   #Assign our transformed data to object
  pivot_wider(names_from = `feature type`,
              values_from = feature_value,)%>%  #Eradicating duplicates by pivot
  rename(age = `1.age`)%>%        #Renaming unfortunate column name
  separate(date,                  #Better visualization be seperating column
           into = c("year", "month"),
           sep="-")%>%
  subset(select = -c(gram, month))%>%  #Removing excessive column
  mutate(age = as.numeric(age)) %>%         #Defining column content as numerical
  mutate(across(.cols = age, ~round(.,digits = 3)))%>% #Keeping to three decimal
  mutate(race =                         #Abbreviating the content in column
           case_when(
             race == "black" ~ "B",
             race == "white" ~ "W",
             race == "none" ~ NA_character_
           ))%>%
  mutate(sex = 
           case_when(
             sex == "female" ~ "0",
             sex == "male" ~ "1",
             sex == "none" ~ NA_character_
           ))%>%
  mutate(blood_neut_pct = if_else(blood_neut_pct <= 35, "Low", "High" )) %>% 
  mutate(blood_cult = 100*blood_cult/max(blood_cult, na.rm = TRUE)) %>% 
  mutate(blood_cult = round(blood_cult, digits = 0))%>% #Keeping to nearest whole number
  mutate(blood_wbc = round(blood_wbc, digits = 1))%>%     #keeping 1 decimal
  mutate(age_agm = age * abm )%>%   #numeric column showing multiplication of age and abm 
  select(id,
         age,
         sex,
         race,
         starts_with("csf"),
         starts_with("blood"),
         everything()) %>%
  full_join(join_data,by = ("id")) #joining the join data file to the mutated tidy_data

#Save file if need be
#write_csv(tidy_data, file = "exam_tidy.csv")
```

```{r Viewing file}

#view(tidy_data) #not advised for large data structures
head(tidy_data)

```



```{r Day 6 Tasks}

#Stratify your data by a categorical column and report min, max, mean and sd of a numeric column.
tidy_data %>% 
  group_by(sex) %>% 
  summarise(max(blood_wbc, na.rm = T), min(blood_wbc, na.rm = T),mean(blood_wbc,na.rm = T),sd(blood_wbc, na.rm =T))
tidy_data %>% 
  group_by(race) %>% 
  summarise(max(blood_wbc, na.rm = T), min(blood_wbc, na.rm = T),mean(blood_wbc,na.rm = T),sd(blood_wbc, na.rm =T))


#Selected column summary (blood_cult, female, etc)
tidy_data %>%       #tidyverse way
  split(.$blood_cult) %>%
  map(summary)

#In order to make it for blood_cult == 0 use filter function (from Akash)
tidy_data %>%         
  filter(blood_cult==0) %>%
  map(summary)

#View as table the relation of blood_gluc grouped by sex
Tabel2 <- 
  tidy_data%>%
  drop_na(sex,
          blood_gluc)%>% #remove missing values to make less messy table
  group_by(sex)%>% #Now select this as the group we check glucose for
  summarize(Lower = min(blood_gluc), #Now we use summarize (NOT summary) to spesify what we look for
            Average = mean(blood_gluc),
            Upper = max(blood_gluc), 
            Difference = max(blood_gluc) - min(blood_gluc)) %>%
  arrange(Average)%>%
  view() 


```



```{r Day 7 }

#Does the glucose level in blood depend on sex? 
#Assuming this task can be solved by plotting information in table2 ?
ggplot(tidy_data,  # define data
       aes(x = as.factor(sex), y = blood_gluc)) +  # define which columns are x and y, 
  geom_boxplot()


#Does the glucose level in CSF (cerebrospinal fluid) depend on race?
csf_gluc_race_grouped <- tidy_data %>%
  filter(race == "B" | race == "W" )%>%
  group_by(race)


#calculate sum of csf_gluc per race
csf_gluc_race_summary <- csf_gluc_race_grouped %>%
  summarise(sum =sum(csf_gluc, na.rm = TRUE))
csf_gluc_race_summary


#Explore your data.
#Explore and comment on the missing variables.
naniar::gg_miss_var((tidy_data))
summary(tidy_data) #R base function
skimr::skim(tidy_data) #Tidyverse way


```

```{r Day 8}

#Analyse the dataset and answer the following questions: (each person chooses one question)
#Is there a difference in the occurrence of the disease by sex? (DYW)
#assign object to select column

tidy_data_4chisq <- tidy_data%>%
  mutate(sex = as.numeric(sex)) %>% #mutating, so that it is same type as abm (R doesn't think like a statistician, so both can be numerical)
  select(abm, sex) #select the variable
#In this test, make sure both variables are same type. 
chisq.test(tidy_data_4chisq$abm, tidy_data_4chisq$sex)


#----------Using regression to find time trend in the occurrence of the disease?#
#Catagorical outcome... makes it difficult
#can view the relation by looking at proportion of disease occurrence pr year
tidy_data%>% 
  filter(!is.na(year)) %>% 
  group_by(year) %>% 
  #  summarise(prop = mean(abm, na.rm=T)) %>% 
  summarise(prop = sum(abm, na.rm = T)) %>% 
  ggplot(aes(x = year, y = prop)) +
  geom_col() 

#Trying linear regression on frequency pr year
Time_trend1 <-
  tidy_data %>%
  group_by(year) %>% 
  summarise(prop = sum(abm, na.rm = T)) %>%
  lm(prop ~ year, data = .) %>%
  broom::tidy()
```





