---
title: "Exam"
author: 'Group_1 : Aditi, Shwesin, Theo, Akash and Daanyaal'
date: "2022-09-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.



#As a group we have devided the tasks between each other
#through each others mistakes we have learnt
#The road is visible in script DYW_Exam_Script
#Now we have produced a final recipe to tidy this specific data
#---------------------- Refined script and steps -----------------------------#


**Packages**

-   [`Reference - ways to load and install packages`](https://statsandr.com/blog/an-efficient-way-to-install-and-load-r-packages/){.uri}


```{r Install and load packages - Alternative 1}

#List of packages
#install.packages(c(tidyverse, ggplot2, here, pacman))
library(tidyverse)
library(here)
library(ggplot2)
library(fs)
library(knitr)
#pacman::p_load(ggplot2, tidyverse, here, fs)

```



** Directory and Environment **
  
  -`Ctrl+Shift+A` to reformat code in the code chunks / R script file
-   `Ctrl+Shift+M` for `%>%`

```{r}
#getwd()  - current directory
#here()   - current directory
#setwd()  - change working directory
#ls()     - files in environment
#remove(ls()) - clears environment - OBS!
#dir()    - files in current directory
#dir_tree(here()) - draws a tree of the current directory
```

** File **
  
  -Loading file into the environment
-   Viewing file + summary stats + missing
-   Data manipulation

```{r Load file, include = FALSE}

#Step 2 - Reading the text files, make sure directory is correct
#Load file into work environment
#Separate columns using delimiter ("tab" - \t)
#Set working directory setwd() if things are not working

here()
setwd("C:/Users/asi006/Downloads/RMED/Advanced_Group_one_project/Data")
nontidy_data <- read_delim("exam_nontidy.txt", delim = "\t")
tidy_data <- nontidy_data
join_data <- read_delim("exam_joindata.txt", delim = "\t")


```

```{r Viewing file,}

#view(tidy_data) #not advised for large data structures
head(tidy_data)
glimpse(tidy_data)

#Provides summary statistics - missing, complete, n and sd
#<https://www.rdocumentation.org/packages/skimr/versions/2.1.4>
skimr::skim(tidy_data)

#Graph showing frequency of missing for each variable
naniar::gg_miss_var((tidy_data))

#column names
colnames(tidy_data)

```


```{r Data wrangling and tidying}

#Day 6 Oppgave
#The Pipe setup of multiple tidy mutations and transformation
#----------------Will yield to the tidiest data -------------------#


tidy_data <- tidy_data %>%   #Assign our transformed data to object
  pivot_wider(names_from = `feature type`,
              values_from = feature_value, ) %>%  #Eradicating duplicates by pivot
  rename(age = `1.age`) %>%        #Renaming unfortunate column name
  separate(date,                  #Better visualization be separating column
           into = c("year", "month"),
           sep = "-") %>%
  mutate(age = as.numeric(age)) %>%         #Defining column content as numerical
  mutate(across(.cols = age, ~ round(., digits = 3))) %>% #Keeping to three decimal across all columns
  mutate(race =                         #Abbreviating the content in column
           case_when(
             race == "black" ~ "B",
             race == "white" ~ "W",
             race == "none" ~ NA_character_
           )) %>%
  mutate(sex =
           case_when(
             sex == "female" ~ "0",
             sex == "male" ~ "1",
             sex == "none" ~ NA_character_
           )) %>%
  mutate(sex = as.factor(sex)) %>%         #Defining column as categorical
  mutate(blood_neut_pct = if_else(blood_neut_pct <= 35, "Low", "High")) %>%
  mutate(blood_cult = 100 * blood_cult / max(blood_cult, na.rm = TRUE)) %>%
  mutate(blood_cult = round(blood_cult, digits = 0)) %>% #Keeping to nearest whole number
  mutate(blood_wbc = round(blood_wbc, digits = 1)) %>%     #keeping 1 decimal
  mutate(age_agm = age * abm) %>%   #numeric column showing multiplication of age and abm
  subset(select = -c(gram, month)) %>%  #Removing excessive column
  select(id,
         age,
         sex,
         race,
         starts_with("csf"),
         starts_with("blood"),
         everything()) %>% #Another way to keep only the columns you want
  full_join(join_data, by = ("id")) %>% #joining the join data file to the mutated tidy_data
  arrange(id) #ordering ID in ascending


# Converts continuous "numeric" into categorical "factor" data
# This methods allows the conversion of several variables into once
factor_cols <-  c("sex", "year", "abm", "race", "blood_neut_pct")
tidy_data[, factor_cols] <-
  lapply(tidy_data[, factor_cols] , factor)


```


```{r Saving file}

#Check path/ directory before saving
#write_csv(tidy_data, file = "exam_tidy.csv")

```


```{r Missing data after tidying the data}

skimr::skim(tidy_data) #Check the complete rate - all rows have an id
naniar::gg_miss_var((tidy_data)) # Graphing the missing data
#The variables "blood_cult" and "csf_cult" have the most missing data (give information on growth of bacteria in the blood and CSF respectively)

```


```{r Day 6 Tasks I}

#frequency/count for each subgroup
tidy_data %>%
  group_by(race, sex) %>%
  count() %>%
  rename(total = n)

#Table
table(tidy_data$race, tidy_data$sex) %>%
  head()


#Explore and comment on the missing variables.
naniar::gg_miss_var((tidy_data))
summary(tidy_data) #R base function
skimr::skim(tidy_data) #Tidyverse way



```

```{r Day 6 Tasks II, warning = FALSE}

# Mean/Min/Max/SD of white blood count in blood (blood_wbc) by sex and race

#blood_wbc_sex <-
tidy_data %>%
  group_by(sex) %>%
  summarise(
    max_blood_wbc = max(blood_wbc, na.rm = T),
    min_blood_wbc = min(blood_wbc, na.rm = T),
    mean_blood_wbc = mean(blood_wbc, na.rm = T),
    sd_blood_wbc = sd(blood_wbc, na.rm = T)
  )

#blood_wbc_race <-
tidy_data %>%
  group_by(race) %>%
  summarise(
    max_blood_wbc = max(blood_wbc, na.rm = T),
    min_blood_wbc = min(blood_wbc, na.rm = T),
    mean_blood_wbc = mean(blood_wbc, na.rm = T),
    sd_blood_wbc = sd(blood_wbc, na.rm = T)
  )

#blood_wbc_race_sex <-
tidy_data %>%
  filter(!is.na(race) &
           !is.na(sex)) %>% #removing cases with missing data for gender and sex
  group_by(race, sex) %>%
  summarise(
    max_blood_wbc = max(blood_wbc, na.rm = T),
    min_blood_wbc = min(blood_wbc, na.rm = T),
    mean_blood_wbc = mean(blood_wbc, na.rm = T),
    sd_blood_wbc = sd(blood_wbc, na.rm = T),
    .groups = 'drop'
  ) # Adding this to remove the warning : https://stackoverflow.com/questions/62140483/how-to-interpret-dplyr-message-summarise-regrouping-output-by-x-override
```

```{r Day 6 Tasks III}

# Mean/Min/Max/SD of csf_prot (Protein in CSF) for people with :

#   blood_cult == 0
tidy_data %>%
  filter(blood_cult == 0) %>%
  summarise(
    max_csf_prot = max(csf_prot, na.rm = T),
    min_csf_prot = min(csf_prot, na.rm = T),
    mean_csf_prot = mean(csf_prot, na.rm = T),
    sd_csf_prot = sd(csf_prot, na.rm = T)
  )

#   females
tidy_data %>%
  filter(sex == 0) %>%
  summarise(
    max_csf_prot = max(csf_prot, na.rm = T),
    min_csf_prot = min(csf_prot, na.rm = T),
    mean_csf_prot = mean(csf_prot, na.rm = T),
    sd_csf_prot = sd(csf_prot, na.rm = T)
  )




#   race black and blood_gluc higher than 120
tidy_data %>%
  filter(blood_gluc >= 120 & race == "B") %>%
  summarise(
    max_csf_prot = max(csf_prot, na.rm = T),
    min_csf_prot = min(csf_prot, na.rm = T),
    mean_csf_prot = mean(csf_prot, na.rm = T),
    sd_csf_prot = sd(csf_prot, na.rm = T)
  )


#   persons older than 45
#Since there is repetition, we decided to create a function to fo this job for us.
data_summary <- function(y) {
  max_ <- max(y, na.rm = T)
  min_ <- min(y, na.rm = T)
  mean_ <- mean(y, na.rm = T)
  sd_ <- sd(y, na.rm = T)
  data_ <- c(max_, min_, mean_, sd_)
  col_ <- c("Maximum", "Minimum", "Mean", "SD")
  df_ <- tibble(col_, data_)
  return(df_)
}

#Although smart, it doesn't seem to work with the pipes.
#Therefore, we created a new data set with the specific filters and ran that as a argument of the function.
tidy_data_age_45 <- tidy_data %>%
  filter(age >= 45)
data_summary(tidy_data_age_45$csf_prot)

```



```{r Day 7 Tasks I}
#    - Are there any correlated measurements?

# We decided to plot a correlation matrix

#https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html

df <- select(tidy_data,-("id")) #removing ID
df <-
  select(df,-("blood_cult")) #removing "blood_cult" since it has a lot of missing values
df <- df[sapply(df, is.numeric)] #making all variables numeric
df <-
  df[complete.cases(df), ] #ensuring only those rows with complete information in all columns are included

cor(df, method = "pearson" , use = "complete.obs") #A correlation matrix for all variables
naniar::gg_miss_var((df)) # Graphing the missing data - none missing

#install.packages("corrplot")
library(corrplot) #can help graphically visualize the correlation plots
corrplot(
  cor(df),
  method = "shade",
  shade.col = NA,
  tl.col = "black",
  tl.srt = 45
)

#only those with p-value < 0.05
testRes = cor.mtest(df, conf.level = 0.95)
corrplot(
  cor(df),
  p.mat = testRes$p,
  method = 'circle',
  type = 'lower',
  insig = 'blank',
  addCoef.col = 'black',
  number.cex = 0.8,
  order = 'AOE',
  diag = FALSE
)

```


```{r Day 7 Task II}
#    - Does the glucose level in blood depend on sex?

tidy_data %>%
  drop_na(sex,
          blood_gluc) %>% #remove missing values to make less messy table
  group_by(sex) %>% #Now select this as the group we check glucose for
  summarize(
    min_blood_gluc = min(blood_gluc),
    mean_blood_gluc = mean(blood_gluc),
    max_blood_gluc = max(blood_gluc),
    sd_blood_gluc = sd(blood_gluc)
  ) %>%
  arrange()
#The average blood glucose level in women is higher than in men


#Graph - Boxplot
tidy_data %>%
  drop_na(sex) %>% #removed missing "sex"
  ggplot(aes(x = as.factor(sex), #to make "sex" categorical
             y = blood_gluc)) +
  geom_boxplot() +
  scale_x_discrete(name = "Gender", labels = c("Female", "Male")) +
  scale_y_continuous(name = "Blood Glucose") +
  theme_minimal() +
  labs(title = "Blood glucose by gender")


```



```{r Day 7 Task III}
#    - Does the glucose level in blood depend on race?

tidy_data %>%
  drop_na(race,
          blood_gluc) %>%
  group_by(race) %>%
  summarize(
    min_blood_gluc = min(blood_gluc),
    mean_blood_gluc = mean(blood_gluc),
    max_blood_gluc = max(blood_gluc),
    sd_blood_gluc = sd(blood_gluc)
  ) %>%
  arrange()
#The average blood glucose level among blacks and whites is similar


#Graph - Boxplot
tidy_data %>%
  drop_na(race) %>%
  ggplot(aes(x = as.factor(race),
             y = blood_gluc)) +
  geom_boxplot() +
  scale_x_discrete(name = "Race", labels = c("Black", "White")) +
  scale_y_continuous(name = "Blood Glucose") +
  theme_minimal() +
  labs(title = "Blood glucose by race")


```


```{r Day 7 Task IV}
#    - Does the glucose level in CSF (cerebrospinal fluid) depend on sex?
# -Does the glucose level in CSF (cerebrospinal fluid) depend on race ?

Gluc_CSF <- tidy_data %>%
  drop_na(race, sex) %>%
  ggplot(aes(x = as.factor(sex),
             y = csf_gluc, fill = sex)) +
  geom_boxplot() +
  scale_x_discrete(name = "", labels = c("Female", "Male")) +
  scale_y_discrete(name = "Glucose in CSF") +
  theme(legend.position = "none")

#Graphing the difference between race and gender on CSF glucose using FACET!
#https://r-graph-gallery.com/264-control-ggplot2-boxplot-colors.html
new_labels <- c("B" = "Black", "W" = "White")
Gluc_CSF + facet_grid(cols = vars(race),
                      labeller = labeller(race = new_labels)) +
  labs(title = "CSF glucose by race")



#Does the glucose level in CSF (cerebrospinal fluid) depend on race?
csf_gluc_race_grouped <- tidy_data %>%
  filter(race == "B" | race == "W") %>%
  group_by(race)
#calculate sum of csf_gluc per race
csf_gluc_race_summary <- csf_gluc_race_grouped %>%
  summarise(sum = sum(csf_gluc, na.rm = TRUE))
csf_gluc_race_summary

```



```{r Day 8 Task I}

#   - Is there a difference in the occurrence of the disease by sex?
tidy_data_4chisq <- tidy_data %>%
  mutate(sex = as.numeric(sex)) %>%
  #mutating, so that it is same type as abm
  #(R doesn't think like a statistician, so both can be numerical)
  select(abm, sex) #select the variable
#In this test, make sure both variables are same class/type
chisq.test(tidy_data_4chisq$abm, tidy_data_4chisq$sex)




```


```{r Day 8 Task II}
#   - Is there a time trend in the occurrence of the disease?

tidy_data %>% 
group_by(year) %>%
  mutate(abm = as.numeric(abm)) %>%
  filter(!is.na(year)) %>%
  #mutate abm into numeric since a factor cannot be summarized
  #doesn't understand statistics
  summarise(sum_abm = sum(abm, na.rm = T)) %>%
  #  table() #For drawing a cross-tab
  ggplot(aes(x = as.factor(year),
             y = sum_abm)) +
  geom_col() +
  scale_x_discrete(name = "Year") +
  scale_y_continuous(name = "No. of patients with acute meningitis cases") +
  theme(legend.position = "none") +
  labs(title = "Number of meningitis over time")


#Trying linear regression on frequency per year
time_trend <-
  tidy_data %>%
  group_by(year) %>%
  mutate(abm = as.numeric(abm)) %>%
  mutate(year = as.numeric(year)) %>%
  
  filter(!is.na(year)) %>%
  summarise(sum_abm = sum(abm, na.rm = T)) %>%
  lm(sum_abm ~ year, data = .) %>%
  broom::tidy()
head(time_trend)

```



```{r Day 8 Task III}

#    - Does the occurrence of the disease depend on age?
#    - Is there a difference in the occurrence of the disease by race?

#https://www.statology.org/r-create-categorical-variable-from-continuous/

tidy_data <- tidy_data %>% 
  mutate(age = as.numeric(age)) 

tidy_data$age_cat <- cut(
  tidy_data$age,
  breaks = c(5, 10, 15, 20,
             25, 30, 35, 40,
             45, 50, 55, 60,
             65, 70, 75, 80, 85),
  labels = c(
    "5",     "10",
    "15",    "20",
    "25",    "30",
    "35",    "40",
    "45",    "50",
    "55",    "60",
    "65",    "70",
    "75",    "80"
  )
)

tidy_data %>%
  mutate(abm = as.numeric(abm)) %>%
  group_by(age_cat, race) %>%
  filter(!is.na(age_cat) & !is.na(race)) %>%
  summarise(sum_abm = sum(abm, na.rm = T)) %>%
  ggplot(aes(
    x = as.factor(age_cat),
    y = sum_abm,
    fill = race
  )) +
  geom_col() +
  scale_x_discrete(name = "Age groups") +
  scale_y_continuous(name = "No. of patients with acute meningitis cases") +
  theme(legend.position = "none") +
  labs(title = "Number of meningitis by age groups") + 
  theme_light()

```
