---
title: "Aditi_Exam_Markdown"
author: "Aditi"
date: "2022-09-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Packages**

-   [`Reference - ways to load and install packages`](https://statsandr.com/blog/an-efficient-way-to-install-and-load-r-packages/){.uri}

```{r Install and load packages - Alternative 1}
#List of packages
packages <- c("tidyverse", "ggplot2", "here", "fs", "knitr")

# Install and load packages
for (p in packages) {
  if (!p %in% installed.packages()) {
    install.packages(p, dependencies = TRUE)
  }
  if (!p %in% .packages()) {
    library(p, character.only = TRUE)
  }
}
```

```{r Install and load packages - Alternative 2}
pacman::p_load(ggplot2, tidyverse, here, fs)
```

**Directory and Environment**

-   `Ctrl+Shift+A` to reformat code in the code chunks/R script file
-   `Ctrl+Shift+M` for `%>%`

```{r }
#getwd()  - current directory
#here()   - current directory
#setwd()  - change working directory
#ls()     - files in environment
#remove(ls()) - clears environment - OBS!
#dir()    - files in current directory
#dir_tree(here()) - draws a tree of the current directory

```

**File**

-   Loading file into the environment
-   Data manipulation

```{r Load file}
#Load file into work environment #Separate columns using delimiter ("tab" - \t)
#setwd("C:/Users/asi006/Downloads/RMED/Advanced_Group_one_project/Data")
read_delim("exam_nontidy.txt", delim = "\t")
tidy_data <- read_delim("exam_nontidy.txt", delim = "\t")
data_untidy <- read_delim("exam_nontidy.txt", delim = "\t")
new_data    <- read_delim("exam_nontidy.txt", delim = "\t")

```

```{r View file}
#view(data_untidy) - not recommended for large datasets
head(data_untidy)
glimpse(data_untidy)

```

```{r Summary statistics}
#columns store values as <chr> which need to be converted to <numeric>
data_untidy %>%
  distinct()
data_untidy %>%
  count(date)
summary(data_untidy)

data_untidy %>%
  summarize(max(blood_wbc, na.rm = T),
            min(blood_wbc, na.rm = T),
            mean(blood_wbc, na.rm = T))

#checking for duplicates
data_untidy %>%
  count(id, sort = TRUE)



#Provides summary statistics - missing, complete, n and sd
#<https://www.rdocumentation.org/packages/skimr/versions/2.1.4>
skimr::skim(data_untidy)

#Graph showing frequency of missing for each variable
naniar::gg_miss_var((data_untidy))



tidy_data%>%
  summarize(Max = max(blood_wbc, na.rm = T),
            Min = min(blood_wbc, na.rm = T),
            Mean = mean(blood_wbc, na.rm = T),
            Sd = sd(blood_wbc, na.rm = T)) 

```

```{r Changing and creating columns}

#Rename column name to "feature_type"
#A good practice to change names which start with numbers or have spaces
data_untidy <-
  new_data %>%
  rename(feature_type = `feature type`,
         age = `1.age`)

#Pivot "feature type" into "sex" and "race" data_untidy \<-
data_untidy <-
  data_untidy %>%
  pivot_wider(names_from  = feature_type,
              values_from = feature_value)

#viewing chosen columns
data_untidy %>%
  select(id, age, date) %>%
  head()


#changing order of columns
#not ending with everything() removes the other columns
data_untidy <-
  data_untidy %>%
  select(id,
         sex,
         race,
         starts_with("csf"),
         starts_with("blood"),
         everything())

#Turning a string column into numeric
data_untidy <-
  data_untidy %>%
  mutate(csf_prot = as.numeric(csf_prot))


#Rounding off decimal points for numeric
data_untidy <-
  data_untidy %>%
  # mutate(across(where(is.numeric), \~ round(., 2)))
  mutate(across(.cols = age, ~ round(., digits = 3))) %>%
  mutate(age = as.numeric(age))

#Creating a new ID column
data_untidy <-
  data_untidy %>%
  mutate(ID_new = 1:n())

#Comparison of two columns
data_untidy %>%
  select(id, ID_new) %>%
  mutate(if_else(as.character(id) == as.character(ID_new), "equal", "different"))

#Removes column "ID_new"
data_untidy <- subset(data_untidy, select = -c(ID_new))


#creating new columns by using a separator
data_untidy <-
  data_untidy %>%
  separate(date, into = c("year", "month"), sep = "-")



#removes dataset from the evironment
#remove(new_data)


#Ifelse and case_when 
data_untidy <-
  data_untidy %>%
  mutate(sex = if_else(sex == "female", "F", "M")) %>%
  mutate(race = case_when(
    race == "black" ~ "black",
    race == "white" ~ "white",
    race == "none" ~ NA_character_
  ))

#Arranging/sorting
data_untidy <-
  data_untidy %>%
  arrange(id)
#  arrange(desc(id))


head(data_untidy)

```

**Merging**

-   Merging files gives me happiness :)

```{r Merging files}
#setwd("C:/Users/asi006/Downloads/RMED/Advanced_Group_one_project/Data")
join_data <- read_delim("exam_joindata.txt", delim = "\t")

join_data <-
  join_data %>%
  arrange(id) %>%
  head()

#merging data - using id column
tidy_data  <-
  full_join(tidy_data, join_data, by = "id") %>%
  arrange(id)

```

```{r Exam question - Day 6}
#Drop columns
tidy_data <-
  tidy_data %>%
  subset(select = -c(gram, year, month))

tidy_data <- 
  tidy_data %>% 
  mutate(n_blood_neut_pct = if_else(blood_neut_pct <= 35, "Low", "High" )) %>% 
  mutate(n_blood_cult = 100*blood_cult/max(blood_cult, na.rm = TRUE)) %>% 
  mutate(n_blood_cult = round(n_blood_cult, digits = 0)) %>% 
  mutate(n_sex = if_else(sex == "F", 1, 0)) %>% 
  mutate(n_age_agm = age * abm ) %>% 
  select(id,
         sex,
         age,
         race,
         starts_with("csf"),
         starts_with("blood"),
         everything()) %>% 
  arrange(id) %>% 
  full_join(join_data, by = "id")

#Data exploration



file <- tidy_data
skimr::skim(file) 
naniar::gg_miss_var(file)
                  

```

**ggPLOT**

```{r Githuv tutorial}

##devtools::install_github("jromanowska/RMED-ggplot-tutorial")
library(tidyverse)
library(patchwork)


consultations_gp <- read_delim(
  "Consultations_cleaned2021-09-03.txt",
  delim = "\t"
)

head(consultations_gp)
## This is publicly available data from SSB on number of consultations at the GP between 2012-2018.


```

```{r Grouping data}

consultations_gp_grouped <- consultations_gp %>% 
  group_by(year, age)
consultations_gp_summary <- consultations_gp_grouped %>% 
  filter(diagnosis != "Alle diagnoser") %>% 
  summarise(total = sum(nConsultations, na.rm = T)) #don't count NA
consultations_gp_summary

```

-   `as.factor(year)` - we want the x-axis to be categorical since the `year` columns are only number and by default would be viewed as a continuous variable.

```{r Graph 1 }
ggplot(consultations_gp_summary, #define data
       aes(x = as.factor(year), y = total)) + #define columns x and y
  geom_point() #define a way to visualize data

```

```{r Graph 2}
plot_sum_consult <- 
  ggplot(consultations_gp_summary, 
         aes(x = as.factor(year), y = total)) +
  geom_point(aes(col = age))

plot_sum_consult

```

```{r Graph 3}
ggplot(consultations_gp_summary,
       aes(x = as.factor(year), y = total)) +
  #geom_col(aes(col = age))
  geom_col(aes(fill = age))

```

```{r Graph 4}

# `group` argument helps draw lines connecting the same categories of points
# run the code below (without the `group` argument) to see how strange the
# result would be
ggplot(consultations_gp_summary,
       aes(x = as.factor(year), y = total)) +
#  aes(group = age)
  geom_line(aes(col = age))

```

```{r Graph 5}

# plotting totals, grouped, non-stacked
ggplot(consultations_gp_summary,
        aes(x = year, y = total)) +
    geom_col(aes(fill = age), position = position_dodge())

```

```{r Graph 6}
# splitting this - checking difference with regard to gender
ggplot(consultations_gp_grouped,
        aes(x = as.factor(year), y = nConsultations)) +
    geom_col(aes(fill = age), position = position_dodge()) +
    facet_wrap(facets = vars(gender))
```

```{r Graph 7 }
# Splitting twice
# splitting again - with regard to age
final.plot <- ggplot(consultations_gp_grouped,
                      aes(x = as.factor(year), y = nConsultations)) +
    geom_col(aes(fill = age), position = position_dodge()) +
    facet_grid(
      cols = vars(gender), 
      rows = vars(diagnosis),
      scales = "free_y")
final.plot

```

```{r Graph 8 }

consultations_gp_0_5 <-   consultations_gp_grouped %>% 
  filter(age == "80-89") 

 a <- ggplot(consultations_gp_0_5,
                        aes(x = as.factor(year), y = nConsultations)) +
  geom_col(aes(fill = gender), position = position_dodge()) +
  facet_grid(rows = vars(diagnosis),
             cols = vars(age),
             scales = "free_y")
a

```

```{r Graph 9 - overlaying a line graph over a bar graph}

my_plot1 <- ggplot(consultations_gp_summary,
        aes(x = as.factor(year), y = total)) +
    geom_col(aes(fill = age), position = position_dodge(width = 0.9))


my_plot1 +
  geom_line(aes(group = age))

```

```{r Graph 10 - adding error bars}
consultations_gp_median <- consultations_gp_grouped %>%
  filter(diagnosis != "Alle diagnoser") %>%
  summarise(
    median = median(nConsultations, na.rm = TRUE),
    stderr = sd(nConsultations, na.rm = TRUE)
  )
consultations_gp_median

my_plot2 <- ggplot(consultations_gp_median,
                   aes(as.factor(year), median)) +
  geom_col(aes(fill = age), position = position_dodge())
my_plot2

#adding error bars
my_plot2 +
  geom_errorbar(aes(
    ymin = (median - stderr),
    ymax = (median + stderr),
    group = age
  ),
  position = position_dodge(width = 0.9))


```

```{r Graph 11 - combining different plots}
plot_sum_consult + my_plot1 # next to each other
final.plot / my_plot1 # on top of each other
(my_plot1 + my_plot2) / (final.plot + plot_sum_consult)

```

```{r Graph 12 - changing the theme and color}
my_plot1 +
  scale_fill_brewer(type = "div", palette = "PRGn")

my_plot1 +
  geom_line(aes(group = age, col = age),
            position = position_dodge(width = 0.9)) +
  scale_fill_brewer(type = "div", palette = "BrBG") +
  scale_color_brewer(type = "div", palette = "PuOr")
```

```{r Graph 13 - adding the axes }
my_plot1 +
  scale_fill_brewer(type = "div", palette = "PRGn") +
  xlab("Year") + # changing the x label
  ylab("Total number of consultations") +
  labs(title = "Number of consultations at general practicioners (GPs)",
       subtitle = "groupped by age",
       caption = "data source: aggregated data from SSB") +
  scale_y_continuous(labels = percent)


my_plot1_nicer <- my_plot1 +
  scale_fill_brewer(type = "div",
                    palette = "PRGn",
                    name = "Age group") + # name of the legend
  xlab("Year") +
  ylab("Total number of  consultations") +
  labs(title = "Number of consultations at general practicioners (GPs)",
       subtitle = "groupped by age",
       caption = "data source: aggregated data from SSB")
my_plot1_nicer


```

```{r Graph 14 - themes and legend}

my_plot1_nicer + theme_minimal()


my_plot1_nicer +
  theme(
    plot.title = element_text(size = 20, # text size in pts
                              face = "bold"),
    axis.title = element_text(size = 16),
    axis.text.x = element_text(angle = 90)
  )

my_plot1_nicer +
  theme(
    legend.position = "bottom"
 )

my_plot1_nicer +
  theme(
      panel.background = element_rect(fill = "salmon")
 )

```

**Day 8**

```{r Day 8}

#install.packages("medicaldata")
#df <- medicaldata::blood_storage

##ANOVA
ANOVAresult <-
  df %>%
  mutate(PVol = log(PVol)) %>%
  aov(PVol ~ RBC.Age.Group, data = .)


ANOVAresult %>% 
  summary()
ANOVAresult %>% 
  broom::tidy()


##T-TEST
t_test <-
  df %>%
  mutate(PVol = log(PVol)) %>%
  t.test(PVol ~ PreopTherapy, data = .) %>% 
  broom::tidy()

t_test_2 <-
  df %>%
  mutate(PVol = log(PVol)) %>%
  group_by(PreopTherapy) %>%
  do(
    t.test(.$PVol, mu = 4) %>%
       broom::tidy()
    )

#MANN-WHITENEY TEST
df %>% 
  wilcox.test(PVol~PreopTherapy, data = .) %>% 
  broom::tidy()


#KRUSKAL-WALLIS TEST
df %>% 
  kruskal.test(PVol~RBC.Age.Group, data = .) %>%
  broom::tidy()


#REGRESSION
df %>% 
  ggplot(aes(x = Age, y = PVol)) + geom_point() + geom_smooth(method = "lm")

df %>% 
  ggplot(aes(x = Age, y = log10(PVol))) + geom_point() + geom_smooth(method = "lm")


lmfit <- df %>%
  lm(PVol ~ Age, data = .) %>%
  broom::tidy()

#REGRESSION - AIC, p-value, r-sqaured
df %>% 
  lm(PVol~Age, data = .) %>%
  broom::glance()

df %>% 
  lm(PVol~Age + PreopPSA, data = .) %>%
  broom::tidy()


```

```{r}
######Data exploration
str(tidy_data)
#Provides the structure of the dataset

summary(tidy_data)
#Frequencies and descriptive statistics - min, Q1, median, mean, Q3, max and NAs

names(tidy_data)
#column names

head(tidy_data, n = 10)
# Default is 6 rows

tidy_data[1:3,1:3]
# The first 3 rows and 3 columns


#####Data missing
sum(is.na(tidy_data))
# Missing in dataset



#####Renaming - oldname = newname
tidy_data <- rename(tidy_data, c("id" = "ID"))



```

```{r Exam}

csf_gluc_race_grouped <- tidy_data %>%
  filter(race == "B" | race == "W" )%>% 
  group_by(race)

#calculate sum of csf_gluc per race
csf_gluc_race_summary <- csf_gluc_race_grouped %>%
  summarise(sum =sum(csf_gluc, na.rm = TRUE))
csf_gluc_race_summary



#ggplot
my_plot1 <- 
                   ggplot(csf_gluc_race_summary,
       aes(x=as.factor(race), y=sum))+
  geom_col(aes(fill = race), position = position_dodge())


my_plot1


my_plot1_nicer <- my_plot1 +
  scale_fill_brewer(type = "div",
                    name = "Glucose in CSF") + # name of the legend
  xlab("Race") +
  ylab("CSF Glucose") +
  labs(title = "CSF Glucose",
       subtitle = "",
       caption = "data source: Group 1")
my_plot1_nicer



df <- tidy_data %>%
  filter(race == "B" | race == "W") %>% 
  t.test(csf_gluc ~ race, data = . ) %>% 
#  broom:: tidy( ) 
  broom:: glance()


tidy_data <-
  tidy_data %>%
  mutate(abm = as.numeric(abm)) %>% 
#   mutate(sex = if_else(sex == "F", 0, 1)) %>% 
    mutate(sex = as.numeric(sex))

  

df <- tidy_data %>%
  select(abm, sex)

df2 <- skimr::skim(df)
df2

library(broom)

df_chi <- chisq.test(df$abm , df$sex) %>% 
    tidy()
df_chi




```
